---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# MOVIE

The **movie** package provides tools for exploring and analyzing summer-themed movies using IMDb data. The package includes a Shiny app for interactive data exploration, functions for filtering and analyzing movie data, and visualizations for insights such as rating trends and genre distributions. The package is designed to help users discover trends, patterns, and statistics related to summer-themed movies.

You can view a comprehensive guide to this package on its [website](https://etc5523-2024.github.io/assignment-4-packages-and-shiny-apps-neharika999/articles/movie.html)

## Installation

You can install the **movie** package directly from GitHub. First, ensure you have the `devtools` package installed, then use `install_github()`:

```{r, eval=FALSE}
# Install devtools if you haven't already
install.packages("devtools")

# Install the movie package from GitHub
devtools::install_github("ETC5523-2024/assignment-4-packages-and-shiny-apps-neharika999/movie")
```

You can install the development version of movie from [GitHub](https://github.com/) with:

## Example

#### Example 1: Filtering Movies by Genre

The `filter_movies_by_genre()` function filters movies by genre. To see only comedy movies:

```{r, eval=FALSE}
# Filtering for comedy movies
filtered_data <- summer_movies %>% filter(genres=="Drama")
head(filtered_data)
```

#### Example: Plotting Average Ratings Over Time

Use `ggplot2` to visualize trends in the filtered data. Here's an example of plotting average ratings over time for a specific genre:

```{r,eval=FALSE}
library(ggplot2)

# Average rating plot for comedy movies over time
ggplot(filtered_data, aes(x = year, y = average_rating)) +
  geom_line(color = "blue") +
  geom_point() +
  labs(title = "Average Ratings of Comedy Movies Over Time",
       x = "Year", y = "Average Rating")

```

## Launching the Shiny App

The package includes an interactive Shiny app that allows users to explore the data through a web interface. You can launch the app using the following command:

```{r,eval=FALSE}
launch_app()
```

The Shiny app provides an easy way to:

-   Select specific genres and filter by rating and year range.

-   View interactive plots and gain insights into summer-themed movies.

## Accessing the Vignette

The package includes a vignette that demonstrates how to use its features. Load the vignette as follows:

```{r,eval=FALSE}
vignette("movie_analysis")
```

## Datasets

### 1.summer_movies

This dataset contains details about summer-themed movies, including:

-   **tconst**: Unique identifier for each movie.

-   **title_type**: Type of the title (e.g., movie, short).

-   **primary_title**: Main title of the movie.

-   **original_title**: Original title of the movie.

-   **year**: Year of release.

-   **runtime_minutes**: Duration of the movie in minutes.

-   **genres**: Genres associated with the movie.

-   **average_rating**: IMDb average rating.

-   **num_votes**: Number of votes the movie received.

### 2. summer_movie_genres

This dataset contains genre information for each movie identified by `tconst`, allowing for filtering and analysis by genre.

## Contributing

If you encounter any issues or would like to contribute to the development of the **movie** package, feel free to create a pull request or submit an issue on GitHub.

## **License**

This package is licensed under the MIT License.
