---
title: "movie"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{movie}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(movie)
```

## Loading and Exploring the Data

The package provides two datasets: `summer_movies` and `summer_movie_genres`.

```{r}
# Loading datasets
data("summer_movies")
data("summer_movie_genres")

# Viewing the first few rows of each dataset
head(summer_movies)
head(summer_movie_genres)
```

### Dataset Descriptions

-   **`summer_movies`**: This dataset contains detailed information about summer-themed movies, including movie titles, release year, runtime, genres, IMDb rating, and the number of votes.

-   **`summer_movie_genres`**: This dataset includes genre information associated with each movie.

## Visualizing Movie Trends

### Average Ratings Over Time

Let's plot the average IMDb ratings of Comedy movies over time to analyze trends.

```{r}
library(ggplot2)
library(dplyr)

# Filter data for Comedy movies
comedy_movies <- summer_movies %>% filter(genres=="comedy")
# Plot average rating over time
ggplot(comedy_movies, aes(x = year, y = average_rating)) +
  geom_line(color = "blue") +
  geom_point() +
  labs(title = "Average Ratings of Comedy Movies Over Time",
       x = "Year", y = "Average Rating")
```

### Distribution of Ratings for a Selected Genre

You can also explore the distribution of IMDb ratings for a specific genre. Here’s how to visualize the distribution of ratings for Comedy movies:

```{r}
ggplot(comedy_movies, aes(x = average_rating)) +
  geom_histogram(binwidth = 0.5, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Ratings for Comedy Movies",
       x = "Average Rating", y = "Frequency")
```

### Popularity (Number of Votes) Over Time

For a genre's popularity, you can examine the number of votes a movie received over time.

```{r}
ggplot(comedy_movies, aes(x = year, y = num_votes)) +
  geom_line(color = "purple") +
  geom_point() +
  labs(title = "Popularity of Comedy Movies Over Time (Number of Votes)",
       x = "Year", y = "Number of Votes")
```

## Using the Shiny App for Interactive Exploration

The `movie` package includes an interactive Shiny app, which provides a user-friendly interface to filter movies by genre, year range, and rating.

## App Layout

The **Explore Summer-Themed Movies** app has two main sections:

1.  **Sidebar Panel**: This is where you apply filters to customize the data displayed in the graphs.

2.  **Main Panel**: This section displays the graphs based on the filters selected in the Sidebar Panel.

### Step-by-Step Guide to Using the App

#### 1. Sidebar Panel

The Sidebar Panel provides filtering options to narrow down the movie data by genre, year range, and minimum rating.

-   **Genre Dropdown**: Select a specific genre (e.g., Comedy, Drama) or choose "All" to include movies from all genres.

    *How to Use*: Select a genre from the dropdown to view only movies within that genre. Choosing "All" shows movies from every genre.

-   **Year Range Slider**: Filter movies by the release year. This slider defaults to the entire range of years in the dataset.

    *How to Use*: Adjust the slider to set the start and end years. For example, setting it from 1990 to 2020 will show only movies released during that period.

-   **Minimum Rating Slider**: Set a minimum IMDb rating to filter for higher-rated movies.

    *How to Use*: Move the slider to set a rating threshold. For instance, a minimum rating of 7 will display only movies rated 7 and above.

-   **Download Button**: Save the filtered dataset to your computer as a CSV file.

------------------------------------------------------------------------

#### 2. Main Panel

The Main Panel displays two visualizations based on the selected filters:

-   **Distribution of Average Ratings**: This histogram shows the distribution of IMDb ratings for movies matching your selected filters. Higher bars indicate more movies with a particular rating range.

-   **Average Rating Over Time**: This line plot shows the trend of average IMDb ratings over time for the selected genre and filters. It helps you observe rating trends and changes over the years.

------------------------------------------------------------------------

### Example Usage

To explore popular Comedy movies from 2000 onwards with high ratings:

1.  **Select Genre**: Choose "Comedy".

2.  **Year Range**: Set the slider to start from 2000.

3.  **Minimum Rating**: Set the rating slider to 7.

4.  **Download Data**: Click "Download Filtered Data" to save this subset.

The graphs will update to show only highly rated Comedy movies from 2000 onwards.

## Summary

The `movie` package provides powerful tools for exploring and analyzing summer-themed movies. You can filter data by genre, visualize trends over time, and use the Shiny app for interactive exploration. This package is ideal for discovering insights about summer-themed movies and understanding how different genres have evolved over the years.

For more details, refer to the function documentation and the Shiny app for further exploration.
